---
title: "BMI"
author: "Clare Horscroft"
date: "04/08/2021"
output: html_document
---

Investigating BMI, weight, and fat measures in relation to number of children

```{r setup}
library(TwoSampleMR)
```

## BMI

```{r }
d <- make_dat("ukb-b-19953", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-19953")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

BMI highly related with number of children, no evidence of reverse causality (i.e. more children does not cause higher BMI)

## Weight

```{r }
d <- make_dat("ukb-b-11842", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-11842")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r}
d <- mv_extract_exposures(c("ukb-b-11842", "ukb-b-19953"))
o <- extract_outcome_data(d$SNP, "ieu-b-4760")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```
Weight effect explained by BMI.


## Hip circumference

```{r }
d <- make_dat("ukb-b-15590", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-15590")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Waist circumference

```{r }
d <- make_dat("ukb-b-9405", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-9405")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r}
d <- mv_extract_exposures(c("ukb-b-15590", "ukb-b-9405", "ukb-b-19953"))
o <- extract_outcome_data(d$SNP, "ieu-b-4760")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Waist/hip circumference effect explained by BMI.

## Body fat

#### Body fat percentage

```{r }
d <- make_dat("ukb-b-8909", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-8909")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Whole body fat mass

```{r }
d <- make_dat("ukb-b-19393", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-19393")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Impedance of whole body

```{r }
d <- make_dat("ukb-b-19921", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-19921")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```



