---
title: "Hair and skin"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating hair and skin-related measures in relation to number of children

```{r setup}
library(TwoSampleMR)
```


## Childhood sunburn occasions

```{r }
d <- make_dat("ukb-b-13246", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-13246")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Ease of skin tanning

```{r }
d <- make_dat("ukb-b-533", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-533")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Use of sun/uv protection 

```{r }
d <- make_dat("ukb-b-7422", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-7422")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Skin colour

```{r }
d <- make_dat("ukb-b-19560", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-19560")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Facial ageing

younger then you are - same - older

```{r }
d <- make_dat("ukb-b-2148", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-2148")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Hair/balding pattern 1

Pattern 1 - full head of hair

```{r }
d <- make_dat("ukb-d-2395_1", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-2395_1")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Hair/balding pattern 3

Pattern 3 - bald spot

```{r }
d <- make_dat("ukb-a-302", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-a-302")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Hair/balding pattern 4

Pattern 4 - totally bald

```{r }
d <- make_dat("ukb-a-303", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-a-303")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Having a full head of hair has a positive effect on number of children, the opposite for other hair patterns.