---
title: "Health"
author: "Clare Horscroft"
date: "06/08/2021"
output: html_document
---

Investigating health measures in relation to number of children

```{r setup}
library(TwoSampleMR)
```

## Overall

#### Overall health rating

```{r }
d <- make_dat("ukb-b-6306", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-6306")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Number of self-reported cancers

```{r }
d <- make_dat("ukb-b-660", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-660")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Eyes

#### Age started wearing glasses or contact lenses

```{r }
d <- make_dat("ukb-b-5801", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-5801")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Spherical power (right)

```{r }
d <- make_dat("ukb-b-19994", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-19994")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Blood

#### Haematocrit percentage

```{r }
d <- make_dat("ukb-d-30030_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30030_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### High light scatter reticulocyte count

```{r }
d <- make_dat("ukb-d-30300_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30300_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### High light scatter reticulocyte percentage

```{r }
d <- make_dat("ukb-d-30290_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30290_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Reticulocyte count

```{r }
d <- make_dat("ukb-d-30250_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30250_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### SHBG 

```{r }
d <- make_dat("ukb-d-30830_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30830_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### IGF-1

```{r }
d <- make_dat("ukb-d-30770_irnt", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-d-30770_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Lungs

#### Forced vital capacity (FVC)

```{r }
d <- make_dat("ukb-b-7953", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-7953")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### FVC

```{r }
d <- make_dat("ieu-b-105", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ieu-b-105")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

