---
title: "Metabolites and proteins"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating metabolites and proteins in relation to number of children

```{r setup}
library(TwoSampleMR)
```


## Metabolites

#### Docosahexaenoic acid

```{r }
d <- make_dat("met-d-DHA", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "met-d-DHA")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Ratio of docosahexaenoic acid to total fatty acids

```{r }
d <- make_dat("met-d-DHA_pct", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "met-d-DHA_pct")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Seems to be some kind of relation between docosahexaenoic acid and number of children.

#### Ratio of omega-3 fatty acids to total fatty acids

```{r }
d <- make_dat("met-d-Omega_3_pct", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "met-d-Omega_3_pct")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Ratio of omega-6 fatty acids to omega-3 fatty acids

```{r }
d <- make_dat("met-d-Omega_6_by_Omega_3", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "met-d-Omega_6_by_Omega_3")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Omega 3 related measures appears to be related to number of children.

```{r}
d <- mv_extract_exposures(c("met-d-DHA_pct", "met-d-Omega_3_pct"))
o <- extract_outcome_data(d$SNP, "ieu-b-4760")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

What does this mean??

## Proteins

Both proteins have very small sample sizes <4000. There may be a relation between these proteins and number of children,

#### Erythrocyte band 7 integral membrane protein

```{r }
d <- make_dat("prot-a-2882", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "prot-a-2882")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Transmembrane protease serine 11D

```{r }
d <- make_dat("prot-a-3024", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "prot-a-3024")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```