---
title: "Sexual behaviour"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating sexual behaviour in relation to number of children

```{r setup}
library(TwoSampleMR)
```

## Age at first live birth

```{r }
d <- make_dat("ukb-b-12405", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-12405")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Age at first live birth has a large effect on the number of children. Interestingly high effect in the reverse (which doesn't make sense temporally) so maybe some form of bias going on.

## Age first had sexual intercourse 

```{r }
d <- make_dat("ukb-b-6591", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-6591")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Very highly related, in BOTH directions..

## Lifetime number of sexual partners 

```{r }
d <- make_dat("ukb-b-4256", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ieu-b-4760", "ukb-b-4256")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

No impact.