---
title: "Activity"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating activity measures in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Jobs

#### Job involves heavy manual or physical work

```{r }
d <- make_dat("ukb-b-2002", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-2002")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Job involves mainly walking or standing

```{r }
d <- make_dat("ukb-b-4461", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-4461")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

These job types seem to have an effect on the number of children, but why?

## Time spent

#### Time spend outdoors in summer

```{r }
d <- make_dat("ukb-b-969", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-969")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Time spent watching television (TV) 

```{r }
d <- make_dat("ukb-b-5192", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-5192")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Usual walking pace

```{r }
d <- make_dat("ukb-b-4711", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-4711")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Weekly usage of mobile phone in last 3 months

```{r }
d <- make_dat("ukb-b-17999", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-17999")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```