---
title: "Health"
author: "Clare Horscroft"
date: "06/08/2021"
output: html_document
---

Investigating health measures in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Overall

#### Overall health rating

```{r }
d <- make_dat("ukb-b-6306", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-6306")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Eyes

#### Age started wearing glasses or contact lenses

```{r }
d <- make_dat("ukb-b-5801", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-5801")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Blood

#### Haematocrit percentage

```{r }
d <- make_dat("ukb-d-30030_irnt", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-d-30030_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### High light scatter reticulocyte count

```{r }
d <- make_dat("ukb-d-30300_irnt", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-d-30300_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### High light scatter reticulocyte percentage

```{r }
d <- make_dat("ukb-d-30290_irnt", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-d-30290_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Reticulocyte count

```{r }
d <- make_dat("ukb-d-30250_irnt", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-d-30250_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### SHBG 

```{r }
d <- make_dat("ukb-d-30830_irnt", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-d-30830_irnt")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### HbA1C

```{r }
d <- make_dat("ieu-b-103", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ieu-b-103")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Lungs

#### Forced vital capacity (FVC)

```{r }
d <- make_dat("ukb-b-7953", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-7953")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### FVC

```{r }
d <- make_dat("ieu-b-105", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ieu-b-105")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Pulse

#### Pulse rate

```{r }
d <- make_dat("ukb-b-18103", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-18103")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Strength

#### Hand grip strength

```{r }
d <- make_dat("ukb-b-10215", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-10215")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Bone density

#### Total body bone mineral density

```{r }
d <- make_dat("ebi-a-GCST005348", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ebi-a-GCST005348")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Most health measures seem to be noisy. Health measures that may be related to number of children: Hand grip strength, FVC, and Age started wearing glasses or contact lenses. Overall health also seems to be related, but there may be reverse causality too.