---
title: "Height"
author: "Clare Horscroft"
date: "06/08/2021"
output: html_document
---

Investigating height in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Sitting height

```{r }
d <- make_dat("ukb-b-16881", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-16881")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Standing height

```{r }
d <- make_dat("ukb-b-10787", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-10787")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Height is highly negatively associated with number of children.

## Is it explained by BMI?

BMI is weight/height^2, so as high BMI -> more children, you'd expect either:
* higher weight to associate with more children (no evidence), or
* higher height to associate with less children (evidence above)
Which is most important, BMI or height?

```{r}
d <- mv_extract_exposures(c("ukb-b-19953", "ukb-b-10787"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Affect of BMI attenuates when height is added to the model.