---
title: "Metabolites and proteins"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating metabolites and proteins in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Fatty acids

#### Ratio of docosahexaenoic acid to total fatty acids

```{r }
d <- make_dat("met-d-DHA_pct", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-d-DHA_pct")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Seems to be some kind of relation between docosahexaenoic acid and number of children.

#### Ratio of omega-3 fatty acids to total fatty acids

```{r }
d <- make_dat("met-d-Omega_3_pct", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-d-Omega_3_pct")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Ratio of omega-6 fatty acids to omega-3 fatty acids

```{r }
d <- make_dat("met-d-Omega_6_by_Omega_3", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-d-Omega_6_by_Omega_3")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Ratio of linoleic acid to total fatty acids

```{r }
d <- make_dat("met-d-LA_pct", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-d-LA_pct")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Average number of double bonds in a fatty acid chain

```{r }
d <- make_dat("met-c-851", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-c-851")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Ratio of bisallylic groups to double bonds

```{r }
d <- make_dat("met-c-844", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-c-844")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Tyrosine

```{r }
d <- make_dat("met-d-Tyr", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "met-d-Tyr")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```