---
title: "SES investigations"
author: "Clare Horscroft"
date: "03/08/2021"
output: html_document
---

Investigating SES measures in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Townsend deprivation index at recruitment

```{r }
d <- make_dat("ukb-b-10011", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-10011")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

No causal relationship, did not appear in list of traits from rms2.

## Average total household income before tax

```{r }
d <- make_dat("ukb-b-7408", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-7408")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Appears to be a causal relationship.

## Years of schooling

```{r }
d <- make_dat("ieu-a-1239", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ieu-a-1239")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Appears to be a causal realtionship.

## Income and Years of Schooling

```{r}
d <- mv_extract_exposures(c("ieu-a-1239", "ukb-b-7408"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```
Income attenuates. Income effect on number of children is mostly explained by years of schooling.

## Age completed full time education

```{r }
d <- make_dat("ukb-a-505", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-a-505")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Qualifications

#### No Qualifications

```{r }
d <- make_dat("ukb-b-17729", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```
```{r }
d <- make_dat("ukb-b-1209", "ukb-b-17729")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### A levels

```{r }
d <- make_dat("ukb-b-11615", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```
```{r }
d <- make_dat("ukb-b-1209", "ukb-b-11615")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Degree

```{r }
d <- make_dat("ukb-b-16489", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```
```{r }
d <- make_dat("ukb-b-1209", "ukb-b-16489")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Can these all be explained using years of schooling?

```{r}
d <- mv_extract_exposures(c("ieu-a-1239", "ukb-b-17729"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)

```
```{r}
d <- mv_extract_exposures(c("ieu-a-1239", "ukb-b-11615"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```
```{r}
d <- mv_extract_exposures(c("ieu-a-1239", "ukb-b-16489"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Qualification effect is explained by years of schooling for no qualifications or A level. However, having a degree is more important.

## Income and degree

```{r}
d <- mv_extract_exposures(c("ukb-b-16489", "ukb-b-7408"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Income effect is explained by having a degree level qualification.

Overall, having a degree explains the difference in number of children for this section.