---
title: "Sexual behaviour"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating sexual behaviour in relation to number of live births

```{r setup}
library(TwoSampleMR)
```

## Age at first live birth

```{r }
d <- make_dat("ukb-b-12405", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-12405")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Age at first live birth has a large effect on the number of children. Interestingly high effect in the reverse (which doesn't make sense temporally) so maybe some form of bias going on.

## Age first had sexual intercourse 

```{r }
d <- make_dat("ukb-b-6591", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-1209", "ukb-b-6591")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Very highly related

Does Age first had sexual intercourse predict Age at first live birth?

```{r }
d <- make_dat("ukb-b-6591", "ukb-b-12405")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Which is more important?

```{r}
d <- mv_extract_exposures(c("ukb-b-12405", "ukb-b-6591"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Age at first live birth is more important, but age first had sexual intercourse is highly predictive. The latter has more associated SNPs.
