---
title: "BMI"
author: "Clare Horscroft"
date: "04/08/2021"
output: html_document
---

Investigating BMI, weight, and fat measures in relation to number of children fathered

```{r setup}
library(TwoSampleMR)
```

## BMI

```{r }
d <- make_dat("ukb-b-19953", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-19953")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

BMI highly related with number of children, no evidence of reverse causality (i.e. more children does not cause higher BMI)

## Weight

```{r }
d <- make_dat("ukb-b-11842", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-11842")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r}
d <- mv_extract_exposures(c("ukb-b-11842", "ukb-b-19953"))
o <- extract_outcome_data(d$SNP, "ukb-b-2227")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Weight effect explained by BMI.

## Waist circumference

```{r }
d <- make_dat("ukb-b-9405", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-9405")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Body fat

#### Whole body fat mass

```{r }
d <- make_dat("ukb-b-19393", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-19393")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

#### Impedance of whole body

```{r }
d <- make_dat("ukb-b-19921", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-19921")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Appendicular lean mass

```{r }
d <- make_dat("ebi-a-GCST90000025", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ebi-a-GCST90000025")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Basal metabolic rate

```{r }
d <- make_dat("ukb-b-16446", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-16446")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r}
d <- mv_extract_exposures(c("ukb-b-16446", "ukb-b-19953"))
o <- extract_outcome_data(d$SNP, "ukb-b-2227")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```