---
title: "Skin"
author: "Clare Horscroft"
date: "05/08/2021"
output: html_document
---

Investigating skin-related measures in relation to number of children fathered

```{r setup}
library(TwoSampleMR)
```

## Childhood sunburn occasions

```{r }
d <- make_dat("ukb-b-13246", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-13246")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Ease of skin tanning

```{r }
d <- make_dat("ukb-b-533", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-533")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Skin colour

```{r }
d <- make_dat("ukb-b-19560", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-19560")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Results show that having more childhood sunburn occasions is negatively related with number of children. Ease of tanning is also negatively related. Having darker skin is positively related.

```{r}
d <- mv_extract_exposures(c("ukb-b-533", "ukb-b-19560"))
o <- extract_outcome_data(d$SNP, "ukb-b-2227")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Ease of skin tanning is more important.