---
title: "Smoking and Alcohol"
author: "Clare Horscroft"
date: "04/08/2021"
output: html_document
---

Investigating smoking and alcohol consumption in relation to number of children fathered

```{r setup}
library(TwoSampleMR)
```

## Alcohol intake frequency

```{r }
d <- make_dat("ukb-b-5779", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-5779")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

No evidence of an effect, perhaps a reverse effect though? Hab=ving more children -> lower alcohol consumption?

## Past tobacco smoking

Note this is coded from 1=Smoke most or all days through to 4=Never smoked

```{r }
d <- make_dat("ukb-b-2134", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-2134")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Never smoked

```{r }
d <- make_dat("ukb-d-20116_0", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-d-20116_0")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Current tabacco smoking

```{r }
d <- make_dat("ukb-b-223", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

```{r }
d <- make_dat("ukb-b-2227", "ukb-b-223")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Smoking and SES

Use degree for SES and Past Smoking

```{r}
d <- mv_extract_exposures(c("ukb-b-16489", "ukb-b-2134"))
o <- extract_outcome_data(d$SNP, "ukb-b-2227")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```
Past tobacco smoking effect on number of children fathered is explained by having a degree.
