---
title: "Fitness is relative"
output:
  html_document:
    df_print: paged
---
  
Check that the 1 in the fitness formula is important.

First set up parameters:
Two genotypes with different effects on phenotype x.
Phenotype x has a fitness effect of 0.05 for every unit above/below the mean.
```{r}
nparents <- 1000000
nkids <- 1000000
b_g1x <- 0.5
b_g2x <- 0.2
g1 <- rbinom(nparents, 2, 0.5)
g2 <- rbinom(nparents, 2, 0.5)
x <- 10 + g1 * b_g1x + g2 * b_g2x + rnorm(nparents)
b_xfitness <- 0.05
```

Define multiple fitness equations with only the leading constant changed:
```{r}
fitness1 <- 1 + b_xfitness * (x - mean(x))
fitness10 <- 10 + b_xfitness * (x - mean(x))
fitness100 <- 100 + b_xfitness * (x - mean(x))
fitness1000 <- 1000 + b_xfitness * (x - mean(x))
```

Calculate the number of children for each fitness equation:
```{r}
calcChildren<-function(fitness){
  child_parent_1 <- sample(1:nparents, nkids, replace=TRUE, prob=fitness)
  child_parent_2 <- sample(1:nparents, nkids, replace=TRUE, prob=fitness)
  
  parents_no_children<-rep(0,nparents)
  tabParents<-table(c(child_parent_1,child_parent_2))
  parents_no_children[as.numeric(unlist(dimnames(tabParents)))]<-tabParents
  return(parents_no_children)
}

parents_no_children_1<-calcChildren(fitness1)
parents_no_children_10<-calcChildren(fitness10)
parents_no_children_100<-calcChildren(fitness100)
parents_no_children_1000<-calcChildren(fitness1000)
```

## How many children by fitness?

### fitness1
```{r}
barplot(table(parents_no_children_1))
summary(lm(parents_no_children_1~fitness1))
plot(0:10,tapply(fitness1[parents_no_children_1<11],parents_no_children_1[parents_no_children_1<11],mean),type="l",xlab="No. Children",ylab="Mean Fitness")
```

### fitness10
```{r}
barplot(table(parents_no_children_10))
summary(lm(parents_no_children_10~fitness10))
plot(0:10,tapply(fitness10[parents_no_children_10<11],parents_no_children_10[parents_no_children_10<11],mean),type="l",xlab="No. Children",ylab="Mean Fitness")
```

### fitness100
```{r}
barplot(table(parents_no_children_100))
summary(lm(parents_no_children_100~fitness100))
plot(0:10,tapply(fitness100[parents_no_children_100<11],parents_no_children_100[parents_no_children_100<11],mean),type="l",xlab="No. Children",ylab="Mean Fitness")
```

### fitness1000
```{r}
barplot(table(parents_no_children_1000))
summary(lm(parents_no_children_1000~fitness1000))
plot(0:10,tapply(fitness1000[parents_no_children_1000<11],parents_no_children_1000[parents_no_children_1000<11],mean),type="l",xlab="No. Children",ylab="Mean Fitness")
```


The linear models and graphs show that for the first fitness equation, the number of children increases as fitness increases. However, for the other equations, the relationship is less clear. 

Therefore, fitness is relative.

Quick way to show this:
```{r}
table(sample(c(1,2,3,4,5),100000,T,prob=c(0.8,0.9,1,1.1,1.2)))
table(sample(c(1,2,3,4,5),100000,T,prob=c(9.8,.9,10,10.1,10.2)))
table(sample(c(1,2,3,4,5),100000,T,prob=c(99.8,99.9,100,100.1,100.2)))
```