---
title: "Intelligence, degree, and number of children"
author: "Clare Horscroft"
date: "18/08/2021"
output: html_document
---

```{r setup}
library(TwoSampleMR)
```

# Number of children

## Effect of intelligence on number of children

```{r}
d <- make_dat("ebi-a-GCST006250", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Effect of degree on number of children

```{r }
d <- make_dat("ukb-b-16489", "ieu-b-4760")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## MVMR for both intelligence and degree on number of children

```{r}
d <- mv_extract_exposures(c("ebi-a-GCST006250", "ukb-b-16489"))
o <- extract_outcome_data(d$SNP, "ieu-b-4760")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Having a degree removes the effect of intelligence.

But does intelligence cause degree? or vice versa?

## Effect of intelligence on degree

```{r }
d <- make_dat("ebi-a-GCST006250", "ukb-b-16489")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Very strong relationship

Now apply Steiger filtering

```{r }
d <- make_dat("ebi-a-GCST006250", "ukb-b-16489")
d <- steiger_filtering(d)
d<-d[d$steiger_dir==TRUE & d$steiger_pval<0.05,]
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Still a strong relationship

## Effect of degree on intelligence

```{r }
d <- make_dat("ukb-b-16489", "ebi-a-GCST006250")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Also a very strong relationship.

Now apply Steiger filtering

```{r }
d <- make_dat("ukb-b-16489","ebi-a-GCST006250")
d <- steiger_filtering(d)
d<-d[d$steiger_dir==TRUE & d$steiger_pval<0.05,]
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

Still a strong relationship.

Therefore, having a degree and intelligence cause each other, but having a degree is more important in terms of number of children.

Does this still hold when split by sex?

# Number of children fathered

## Effect of intelligence on number of children fathered

```{r }
d <- make_dat("ebi-a-GCST006250", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Effect of degree on number of children fathered

```{r }
d <- make_dat("ukb-b-16489", "ukb-b-2227")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## MVMR for both intelligence and degree on number of children fathered

```{r}
d <- mv_extract_exposures(c("ebi-a-GCST006250", "ukb-b-16489"))
o <- extract_outcome_data(d$SNP, "ukb-b-2227")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Having a degree removes the effect of intelligence.

# Number of live births

## Effect of intelligence on number of live births

```{r }
d <- make_dat("ebi-a-GCST006250", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## Effect of degree on number of live births 

```{r }
d <- make_dat("ukb-b-16489", "ukb-b-1209")
d_mr<-mr(d)
d_mr
mr_scatter_plot(d_mr,d)
```

## MVMR for both intelligence and degree on number of live births

```{r}
d <- mv_extract_exposures(c("ebi-a-GCST006250", "ukb-b-16489"))
o <- extract_outcome_data(d$SNP, "ukb-b-1209")
d <- mv_harmonise_data(d, o)
mv_multiple(d)
```

Having a degree removes the effect of intelligence.