---
title: "MR with grandchildren"
author: "Clare Horscroft"
date: "10/05/2021"
output: html_document
---

Investigations into grandchildren

Use the first generation from the multigen simulations to get grandchildren information

```{r}
require(TwoSampleMR)
require(jsonlite)
config <- read_json("~/config.json")
mr_sel_path <- config$mr_sel_path
```

Get data
```{r}
dat<-read.table(paste0(mr_sel_path,"/Simulations/children/multigen/selStop_sim_1/dat_gc_0.txt"),header = TRUE)
hist(dat$grandchildren)
```
MR analysis for children and grandchildren from all sims over all generations
```{r}
b_c_mat<-matrix(NA,100,19)
se_c_mat<-matrix(NA,100,19)
b_gc_mat<-matrix(NA,100,19)
se_gc_mat<-matrix(NA,100,19)
for (s in 1:100){
  setwd(paste0(mr_sel_path,"/Simulations/children/multigen/selStop_sim_",s))
  for (g in 0:18){
    dat<-read.table(paste0("dat_gc_",g,".txt"),header = TRUE)
    mod1<-summary(lm(dat$pheno~dat$geno))
    bgx<-mod1$coefficients[2,1]
    segx<-mod1$coefficients[2,2]
    mod2<-summary(glm(dat$children~dat$geno,family=poisson("log")))
    bgy<-mod2$coefficients[2,1]
    segy<-mod2$coefficients[2,2]
    mrw<-mr_wald_ratio(bgx,bgy,segx,segy)
    mod3<-summary(glm(dat$grandchildren~dat$geno,family=poisson("log")))
    bgygc<-mod3$coefficients[2,1]
    segygc<-mod3$coefficients[2,2]
    mrwgc<-mr_wald_ratio(bgx,bgygc,segx,segygc)
    b_c_mat[s,g+1]<-mrw$b
    se_c_mat[s,g+1]<-mrw$se
    b_gc_mat[s,g+1]<-mrwgc$b
    se_gc_mat[s,g+1]<-mrwgc$se
  }
}
```

box plot for children
```{r}
boxplot(b_c_mat,xlab="Generations",ylab="Estimated selection strength")
abline(h=0.05,lty=2,col="red")
abline(h=0,lty=2,col="blue")
```

box plot for grandchildren
```{r}
boxplot(b_gc_mat,xlab="Generations",ylab="Estimated selection strength")
abline(h=0.05*3/2,lty=2,col="green")
abline(h=0.05,lty=2,col="red")
abline(h=0,lty=2,col="blue")
```

```{r}
apply(b_gc_mat,2,mean)
```

The selection strength is estimated as 1.5 times the actual selection strength when using grandchildren. This is because the beneficial allele is passed down through multiple generations.

If the selection strength is b, and the beneficial allele is a:

In generation 0 (grandparents):

AA: Fitness = 1

Aa: Fitness = 1 + b

aa: Fitness = 1 + 2b


Difference between the groups is b. This is the additional fitness for children for each allele.

The fitness in terms of grandchildren per allele also depends on the likelihood of passing on the beneficial allele:

AA: Fitness = 1 + 0 (no chance of passing on allele a) = 1

Aa: Fitness = 1 + b + 0.5*b (half a chance of passing on allele a) = 1 + 1.5b

aa: Fitness = 1 + 2b + b (Definitely will pass on allele a) = 1 + 3b


Difference between the groups is 1.5b. This is the additional fitness for grandchildren for each allele.