---
title: "Multiple generations with age structure"
author: "Clare Horscroft"
date: "05/05/2021"
output: html_document
---

Create a population with an age structure. Ages represent decades. There will always be an equal number of people in each age bracket. 7 is the max age. Reproduction only happens in ages 2-4.

# Selection strength = 0.05

```{r}
require(TwoSampleMR)
```

Initial population set up:
```{r}
N<-100000
age<-rep(c(0:7),each=N/8)
genotype<-sample(c(0,1,2),N,T)
phenotype<-genotype+rnorm(N,0,1)
fitness<-exp(1+phenotype*0.05+rnorm(N,0,0.01))
nochildren<-rep(0,N)
```

10 decade burn-in
```{r}
for (dec in 1:10){
  #only ages 2-4 can have children
  child_fit<-fitness
  child_fit[age<=1 | age >=5]<-0
  parents<-sample(1:N,N/4,T,prob=child_fit)
  tabParents <- table(parents)
  newchildren<-rep(0,N)
  newchildren[as.numeric(unlist(dimnames(tabParents)))] <-tabParents
  nochildren<-nochildren+newchildren
  
  # get the genotype of the children based on the parents' genotypes
  newgeno<-rep(0,N/8)
  for(i in 1:(N/8)){
    p1<-genotype[parents[i*2-1]]
    p2<-genotype[parents[i*2]]
    if (p1==2){
      newgeno[i]<-newgeno[i]+1
    } else if (p1==1){
      newgeno[i]<-newgeno[i]+sample(c(0,1),1)
    }
    if (p2==2){
      newgeno[i]<-newgeno[i]+1
    } else if (p2==1){
      newgeno[i]<-newgeno[i]+sample(c(0,1),1)
    }
  }
  # get phenotype and fitness for the child given the genotype
  newpheno<-newgeno+rnorm(N/8,0,1)
  newfitness<-exp(1+newpheno*0.05+rnorm(N/8,0,0.01))
  
  ## iterate population 
  # age stays the same as the same amount of people are born as die, and all other information shifts along
  genotype<-c(newgeno,genotype[-c((N-N/8+1):N)])
  phenotype<-c(newpheno,phenotype[-c((N-N/8+1):N)])
  fitness<-c(newfitness,fitness[-c((N-N/8+1):N)])
  nochildren<-c(rep(0,N/8),nochildren[-c((N-N/8+1):N)])
}
```

Overall effect estimate
```{r}
mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]))
bgx<-mod1$coefficients[2,1]
segx<-mod1$coefficients[2,2]
mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3],family=poisson("log")))
bgy<-mod2$coefficients[2,1]
segy<-mod2$coefficients[2,2]
wr<-mr_wald_ratio(bgx,bgy,segx,segy)
wr
```

split by age
```{r}
effectbyage<-rep(0,5)
for(a in 3:7){
  mod1<-summary(lm(phenotype[age==a]~genotype[age==a]))
  bgx<-mod1$coefficients[2,1]
  segx<-mod1$coefficients[2,2]
  mod2<-summary(glm(nochildren[age==a]~genotype[age==a],family=poisson("log")))
  bgy<-mod2$coefficients[2,1]
  segy<-mod2$coefficients[2,2]
  wr<-mr_wald_ratio(bgx,bgy,segx,segy)
  effectbyage[a-2]<-wr$b
}
effectbyage
```

Adjusting for age
```{r}
mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]+age[age>=3]))
bgx<-mod1$coefficients[2,1]
segx<-mod1$coefficients[2,2]
mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3]+age[age>=3],family=poisson("log")))
bgy<-mod2$coefficients[2,1]
segy<-mod2$coefficients[2,2]
wr<-mr_wald_ratio(bgx,bgy,segx,segy)
wr
```

# Selection pressures removed

What happens if the selection pressure is removed after 10 more decades?

```{r}
oee_b<-rep(0,25)
oee_se<-rep(0,25)
sba_b<-matrix(0,5,25)
sba_se<-matrix(0,5,25)
afa_b<-rep(0,25)
afa_se<-rep(0,25)

for (dec in 1:25){
  #only ages 2-4 can have children
  child_fit<-fitness
  child_fit[age<=1 | age >=5]<-0
  parents<-sample(1:N,N/4,T,prob=child_fit)
  tabParents <- table(parents)
  newchildren<-rep(0,N)
  newchildren[as.numeric(unlist(dimnames(tabParents)))] <-tabParents
  nochildren<-nochildren+newchildren
  
  # get the genotype of the children based on the parents' genotypes
  newgeno<-rep(0,N/8)
  for(i in 1:(N/8)){
    p1<-genotype[parents[i*2-1]]
    p2<-genotype[parents[i*2]]
    if (p1==2){
      newgeno[i]<-newgeno[i]+1
    } else if (p1==1){
      newgeno[i]<-newgeno[i]+sample(c(0,1),1)
    }
    if (p2==2){
      newgeno[i]<-newgeno[i]+1
    } else if (p2==1){
      newgeno[i]<-newgeno[i]+sample(c(0,1),1)
    }
  }
  # get phenotype and fitness for the child given the genotype
  newpheno<-newgeno+rnorm(N/8,0,1)
  if(dec<10){ # At generation 10, selection pressure stops
  newfitness<-exp(1+newpheno*0.05+rnorm(N/8,0,0.01))
  } else {
    newfitness<-exp(1+rnorm(N/8,0,0.01))
  }
  
  # Selection pressure stops
  if (dec==10){
      fitness<-exp(log(fitness)-phenotype*0.05)
  }
  ## iterate population 
  # age stays the same as the same amount of people are born as die, and all other information shifts along
  genotype<-c(newgeno,genotype[-c((N-N/8+1):N)])
  phenotype<-c(newpheno,phenotype[-c((N-N/8+1):N)])
  fitness<-c(newfitness,fitness[-c((N-N/8+1):N)])
  nochildren<-c(rep(0,N/8),nochildren[-c((N-N/8+1):N)])
  
  # Overall effect estimate
  mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]))
  bgx<-mod1$coefficients[2,1]
  segx<-mod1$coefficients[2,2]
  mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3],family=poisson("log")))
  bgy<-mod2$coefficients[2,1]
  segy<-mod2$coefficients[2,2]
  wr<-mr_wald_ratio(bgx,bgy,segx,segy)
  oee_b[dec]<-wr$b
  oee_se[dec]<-wr$se

  # Split by age
  for(a in 3:7){
    mod1<-summary(lm(phenotype[age==a]~genotype[age==a]))
    bgx<-mod1$coefficients[2,1]
    segx<-mod1$coefficients[2,2]
    mod2<-summary(glm(nochildren[age==a]~genotype[age==a],family=poisson("log")))
    bgy<-mod2$coefficients[2,1]
    segy<-mod2$coefficients[2,2]
    wr<-mr_wald_ratio(bgx,bgy,segx,segy)
    sba_b[a-2,dec]<-wr$b
    sba_se[a-2,dec]<-wr$se
  }
  
  # Adjusting for age
  mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]+age[age>=3]))
  bgx<-mod1$coefficients[2,1]
  segx<-mod1$coefficients[2,2]
  mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3]+age[age>=3],family=poisson("log")))
  bgy<-mod2$coefficients[2,1]
  segy<-mod2$coefficients[2,2]
  wr<-mr_wald_ratio(bgx,bgy,segx,segy)
  afa_b[dec]<-wr$b
  afa_se[dec]<-wr$se
}
```

Plot selection estimates
```{r}
plot(1:25,oee_b,type="n",ylim=c(min(oee_b-1.96*oee_se,afa_b-1.96*afa_se),max(oee_b+1.96*oee_se,afa_b+1.96*afa_se)),xaxs="i",xlab="Decades",ylab="Selection strength")
polygon(x=c(1:25,25:1),y=c(afa_b+afa_se*1.96,rev(afa_b-afa_se*1.96)),col="lightblue",density = 10,border="lightblue",lty=2)
polygon(x=c(1:25,25:1),y=c(oee_b+oee_se*1.96,rev(oee_b-oee_se*1.96)),col="lightpink",density = 10,border="lightpink",angle=-45)
abline(h=0.05)
abline(h=0)
lines(1:25,oee_b,col="red")
lines(1:25,afa_b,col="blue",lty=2)
legend("topright",legend=c("Overall estimate","Age adjusted"),lty=c(1,2),col=c("red","blue"))
```

Plot split by age
```{r}
plot(1:25,sba_b[1,],type="l",col=1)
lines(1:25,sba_b[2,],col=2)
lines(1:25,sba_b[3,],col=3)
lines(1:25,sba_b[4,],col=4)
lines(1:25,sba_b[5,],col=5)
legend("topright",legend=c(3,4,5,6,7),col=c(1:5),lty=1)

```

# Repeat simulation 100 times

```{r}
  
oee_b<-matrix(0,100,25)
oee_se<-matrix(0,100,25)
sba_b<-list(matrix(0,100,25),matrix(0,100,25),matrix(0,100,25),matrix(0,100,25),matrix(0,100,25))
sba_se<-list(matrix(0,100,25),matrix(0,100,25),matrix(0,100,25),matrix(0,100,25),matrix(0,100,25))
afa_b<-matrix(0,100,25)
afa_se<-matrix(0,100,25)

for (sim in 1:100){
  N<-100000
  age<-rep(c(0:7),each=N/8)
  genotype<-sample(c(0,1,2),N,T)
  phenotype<-genotype+rnorm(N,0,1)
  fitness<-exp(1+phenotype*0.05+rnorm(N,0,0.01))
  nochildren<-rep(0,N)
  
  # 10 decade burn-in
  
  for (dec in 1:10){
    #only ages 2-4 can have children
    child_fit<-fitness
    child_fit[age<=1 | age >=5]<-0
    parents<-sample(1:N,N/4,T,prob=child_fit)
    tabParents <- table(parents)
    newchildren<-rep(0,N)
    newchildren[as.numeric(unlist(dimnames(tabParents)))] <-tabParents
    nochildren<-nochildren+newchildren
    
    # get the genotype of the children based on the parents' genotypes
    newgeno<-rep(0,N/8)
    for(i in 1:(N/8)){
      p1<-genotype[parents[i*2-1]]
      p2<-genotype[parents[i*2]]
      if (p1==2){
        newgeno[i]<-newgeno[i]+1
      } else if (p1==1){
        newgeno[i]<-newgeno[i]+sample(c(0,1),1)
      }
      if (p2==2){
        newgeno[i]<-newgeno[i]+1
      } else if (p2==1){
        newgeno[i]<-newgeno[i]+sample(c(0,1),1)
      }
    }
    # get phenotype and fitness for the child given the genotype
    newpheno<-newgeno+rnorm(N/8,0,1)
    newfitness<-exp(1+newpheno*0.05+rnorm(N/8,0,0.01))
    
    ## iterate population 
    # age stays the same as the same amount of people are born as die, and all other information shifts along
    genotype<-c(newgeno,genotype[-c((N-N/8+1):N)])
    phenotype<-c(newpheno,phenotype[-c((N-N/8+1):N)])
    fitness<-c(newfitness,fitness[-c((N-N/8+1):N)])
    nochildren<-c(rep(0,N/8),nochildren[-c((N-N/8+1):N)])
  }
  
  # next 25 decades
  
  for (dec in 1:25){
    #only ages 2-4 can have children
    child_fit<-fitness
    child_fit[age<=1 | age >=5]<-0
    parents<-sample(1:N,N/4,T,prob=child_fit)
    tabParents <- table(parents)
    newchildren<-rep(0,N)
    newchildren[as.numeric(unlist(dimnames(tabParents)))] <-tabParents
    nochildren<-nochildren+newchildren
    
    # get the genotype of the children based on the parents' genotypes
    newgeno<-rep(0,N/8)
    for(i in 1:(N/8)){
      p1<-genotype[parents[i*2-1]]
      p2<-genotype[parents[i*2]]
      if (p1==2){
        newgeno[i]<-newgeno[i]+1
      } else if (p1==1){
        newgeno[i]<-newgeno[i]+sample(c(0,1),1)
      }
      if (p2==2){
        newgeno[i]<-newgeno[i]+1
      } else if (p2==1){
        newgeno[i]<-newgeno[i]+sample(c(0,1),1)
      }
    }
    # get phenotype and fitness for the child given the genotype
    newpheno<-newgeno+rnorm(N/8,0,1)
    if(dec<10){ # At generation 10, selection pressure stops
    newfitness<-exp(1+newpheno*0.05+rnorm(N/8,0,0.01))
    } else {
      newfitness<-exp(1+rnorm(N/8,0,0.01))
    }
    
    # Selection pressure stops
    if (dec==10){
        fitness<-exp(log(fitness)-phenotype*0.05)
    }
    ## iterate population 
    # age stays the same as the same amount of people are born as die, and all other information shifts along
    genotype<-c(newgeno,genotype[-c((N-N/8+1):N)])
    phenotype<-c(newpheno,phenotype[-c((N-N/8+1):N)])
    fitness<-c(newfitness,fitness[-c((N-N/8+1):N)])
    nochildren<-c(rep(0,N/8),nochildren[-c((N-N/8+1):N)])
    
    # Overall effect estimate
    mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]))
    bgx<-mod1$coefficients[2,1]
    segx<-mod1$coefficients[2,2]
    mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3],family=poisson("log")))
    bgy<-mod2$coefficients[2,1]
    segy<-mod2$coefficients[2,2]
    wr<-mr_wald_ratio(bgx,bgy,segx,segy)
    oee_b[sim,dec]<-wr$b
    oee_se[sim,dec]<-wr$se
  
    # Split by age
    for(a in 3:7){
      mod1<-summary(lm(phenotype[age==a]~genotype[age==a]))
      bgx<-mod1$coefficients[2,1]
      segx<-mod1$coefficients[2,2]
      mod2<-summary(glm(nochildren[age==a]~genotype[age==a],family=poisson("log")))
      bgy<-mod2$coefficients[2,1]
      segy<-mod2$coefficients[2,2]
      wr<-mr_wald_ratio(bgx,bgy,segx,segy)
      sba_b[[a-2]][sim,dec]<-wr$b
      sba_se[[a-2]][sim,dec]<-wr$se
    }
    
    # Adjusting for age
    mod1<-summary(lm(phenotype[age>=3]~genotype[age>=3]+age[age>=3]))
    bgx<-mod1$coefficients[2,1]
    segx<-mod1$coefficients[2,2]
    mod2<-summary(glm(nochildren[age>=3]~genotype[age>=3]+age[age>=3],family=poisson("log")))
    bgy<-mod2$coefficients[2,1]
    segy<-mod2$coefficients[2,2]
    wr<-mr_wald_ratio(bgx,bgy,segx,segy)
    afa_b[sim,dec]<-wr$b
    afa_se[sim,dec]<-wr$se
  }
}
```

Plot overall effect estimate and adjusted by age
```{r}
box<-data.frame(b=c(oee_b[1,],afa_b[1,]),sim=1,dec=rep(1:25,2),dat=rep(c("oee","afa"),each=25))
for (i in 2:100){
  box<-rbind(box,data.frame(b=c(oee_b[i,],afa_b[i,]),sim=i,dec=rep(1:25,2),dat=rep(c("oee","afa"),each=25)))
}
boxplot(b~dat+dec,data = box,col=c("red","blue"),xaxt="n")
axis(side = 1, at = seq(1.5,49.5,2),labels = c(1:25))
legend("topright", fill = c("red","blue"), legend = c("oee","afa"))
```
Plot split by age
```{r}
box<-data.frame(b=c(sba_b[[1]][1,],sba_b[[2]][1,],sba_b[[3]][1,],sba_b[[4]][1,],sba_b[[5]][1,]),sim=1,dec=rep(1:25,5),age=rep(c(3:7),each=25))
for (i in 2:100){
  box<-rbind(box,data.frame(b=c(sba_b[[1]][i,],sba_b[[2]][i,],sba_b[[3]][i,],sba_b[[4]][i,],sba_b[[5]][i,]),sim=i,dec=rep(1:25,5),age=rep(c(3:7),each=25)))
}
boxplot(b~age+dec,data = box,col=c("grey0","grey25","grey50","grey75","grey100"),xaxt="n")
axis(side = 1, at = seq(3,123,5),labels = c(1:25))
legend("topright", fill = c("grey0","grey25","grey50","grey75","grey100"), legend = c(3:7))

# dec 8-17
box_817<-box[box$dec>=8 & box$dec<=17,]
boxplot(b~age+dec,data = box_817,col=c("grey0","grey25","grey50","grey75","grey100"),xaxt="n")
axis(side = 1, at = seq(3,48,5),labels = c(8:17))
legend("topright", fill = c("grey0","grey25","grey50","grey75","grey100"), legend = c(3:7))
```

Recovering selection

Using the decades data should mean that the original selection strength is recoverable. For example, in decade 13. The selection strength should be around a third for age 6 and two-thirds for age 7
```{r}
#age 7
mean(sba_b[[5]][,13])*3/2
#age 6
mean(sba_b[[4]][,13])*3

```