---
title: "Analysis when there are multiple genes and phenotypes afftecting fitness"
author: "Clare Horscroft"
date: "16/03/2021"
output: html_document
---

Load packages
```{r}
require(systemfit)
require(TwoSampleMR)
require(ggdag)
require(dagitty)
require(jsonlite)
```

Get file path
```{r}
config <- read_json("~/config.json")
mr_sel_path <- config$mr_sel_path
```

# Two phenotypes and seven genes

All five genes contribute to the main phenotype, but one gene affects a second phenotype.
Both phenotypes impact fitness.
For this scenario the DAG looks like this:
```{r}
dag <- dagitty('dag{
              SNP1 [pos="0,7"]
              SNP2 [pos="0,6"]
              SNP3 [pos="0,5"]  
              SNP4 [pos="0,4"]
              SNP5 [pos="0,3"]
              SNP6 [pos="0,2"]
              SNP7 [pos="0,1"]
              Trait1 [pos="1,5"]
              Trait2 [pos="1,3"]
              Fitness  [pos="2,4"]
              Confounder [pos="1.5,7"]
              SNP1 -> Trait1 -> Fitness
              SNP2 -> Trait1 -> Fitness
              SNP3 -> Trait1 -> Fitness
              SNP4 -> Trait1 -> Fitness
              SNP4 -> Trait2 -> Fitness
              SNP5 -> Trait2 -> Fitness
              SNP6 -> Trait2 -> Fitness
              SNP7 -> Trait2 -> Fitness
              Trait1 <- Confounder -> Fitness
}')
ggdag(dag, layout = "circle",node_size = 20, text_size = 2.7) +
  theme_void()
```

```{r}
dat<-read.table(paste0(mr_sel_path,"/Simulations/children/two_pheno_7_gene/dat.txt"),header = TRUE)
```

## MR analysis

Using the systemfit method on fitness
```{r}
sf1<-systemfit(log(fit)~pheno1+pheno2,inst = ~geno1+geno2+geno3+geno4+geno5+geno6+geno7,method = "2SLS",data=dat)
summary(sf1)
```

### MR for phenotype 1

Using the mr function
```{r}
exposure<-t(sapply(2:8,function(x){temp<-summary(lm(dat$pheno1~dat[,x]));c(temp$coefficients[2,c(1,2)],fstatistic=temp$fstatistic[1],rsquared=temp$r.squared)}))
outcome<-t(sapply(2:8,function(x)summary(lm(log(dat$fit)~dat[,x]))$coefficients[2,c(1,2)]))
dat_pheno1<-data.frame(SNP=1:7,id.exposure="pheno1",exposure="pheno1",id.outcome="fit",outcome="fit",beta.exposure=exposure[,1],se.exposure=exposure[,2],beta.outcome=outcome[,1],se.outcome=outcome[,2],mr_keep=TRUE)

#only keep the four snps associated with phenotype1
dat_pheno1<-dat_pheno1[1:4,]

dat_pheno1$beta.outcome[dat_pheno1$beta.exposure<0]<--dat_pheno1$beta.outcome[dat_pheno1$beta.exposure<0]
dat_pheno1$beta.exposure[dat_pheno1$beta.exposure<0]<--dat_pheno1$beta.exposure[dat_pheno1$beta.exposure<0]
mr_results <- mr(dat_pheno1, method_list=c("mr_ivw","mr_egger_regression","mr_weighted_median", "mr_weighted_mode"))
mr_results
```


Sensitivity analysis
```{r}
mr_scatter_plot(mr_results, dat_pheno1)
mr_forest_plot(mr_singlesnp(dat_pheno1))
```


### Phenotype 1 with number of children as the outcome

```{r}
exposure<-t(sapply(2:8,function(x){temp<-summary(lm(dat$pheno1~dat[,x]));c(temp$coefficients[2,c(1,2)],fstatistic=temp$fstatistic[1],rsquared=temp$r.squared)}))
outcome<-t(sapply(2:8,function(x)summary(glm(dat$children~dat[,x],family = poisson()))$coefficients[2,c(1,2)]))
dat_pheno1<-data.frame(SNP=1:7,id.exposure="pheno1",exposure="pheno1",id.outcome="child",outcome="child",beta.exposure=exposure[,1],se.exposure=exposure[,2],beta.outcome=outcome[,1],se.outcome=outcome[,2],mr_keep=TRUE)

#only keep the four snps associated with phenotype1
dat_pheno1<-dat_pheno1[1:4,]

dat_pheno1$beta.outcome[dat_pheno1$beta.exposure<0]<--dat_pheno1$beta.outcome[dat_pheno1$beta.exposure<0]
dat_pheno1$beta.exposure[dat_pheno1$beta.exposure<0]<--dat_pheno1$beta.exposure[dat_pheno1$beta.exposure<0]
mr_results <- mr(dat_pheno1, method_list=c("mr_ivw","mr_egger_regression","mr_weighted_median", "mr_weighted_mode"))
mr_results
```

```{r}
mr_heterogeneity(dat_pheno1)
mr_scatter_plot(mr_results, dat_pheno1)
mr_forest_plot(mr_singlesnp(dat_pheno1))
```

### MR for phenotype 2

Using the mr function
```{r}
exposure<-t(sapply(2:8,function(x){temp<-summary(lm(dat$pheno2~dat[,x]));c(temp$coefficients[2,c(1,2)],fstatistic=temp$fstatistic[1],rsquared=temp$r.squared)}))
outcome<-t(sapply(2:8,function(x)summary(lm(log(dat$fit)~dat[,x]))$coefficients[2,c(1,2)]))
dat_pheno2<-data.frame(SNP=1:7,id.exposure="pheno2",exposure="pheno2",id.outcome="fit",outcome="fit",beta.exposure=exposure[,1],se.exposure=exposure[,2],beta.outcome=outcome[,1],se.outcome=outcome[,2],mr_keep=TRUE)

#only keep the four snps associated with phenotype1
dat_pheno2<-dat_pheno2[4:7,]

dat_pheno2$beta.outcome[dat_pheno2$beta.exposure<0]<--dat_pheno2$beta.outcome[dat_pheno2$beta.exposure<0]
dat_pheno2$beta.exposure[dat_pheno2$beta.exposure<0]<--dat_pheno2$beta.exposure[dat_pheno2$beta.exposure<0]
mr_results <- mr(dat_pheno2, method_list=c("mr_ivw","mr_egger_regression","mr_weighted_median", "mr_weighted_mode"))
mr_results
```


Sensitivity analysis
```{r}
mr_heterogeneity(dat_pheno2)
mr_scatter_plot(mr_results, dat_pheno2)
mr_forest_plot(mr_singlesnp(dat_pheno2))
```


### Phenotype 2 with number of children as the outcome

```{r}
exposure<-t(sapply(2:8,function(x){temp<-summary(lm(dat$pheno2~dat[,x]));c(temp$coefficients[2,c(1,2)],fstatistic=temp$fstatistic[1],rsquared=temp$r.squared)}))
outcome<-t(sapply(2:8,function(x)summary(glm(dat$children~dat[,x],family = poisson()))$coefficients[2,c(1,2)]))
dat_pheno2<-data.frame(SNP=1:7,id.exposure="pheno2",exposure="pheno2",id.outcome="child",outcome="child",beta.exposure=exposure[,1],se.exposure=exposure[,2],beta.outcome=outcome[,1],se.outcome=outcome[,2],mr_keep=TRUE)

#only keep the four snps associated with phenotype2
dat_pheno2<-dat_pheno2[4:7,]

dat_pheno2$beta.outcome[dat_pheno2$beta.exposure<0]<--dat_pheno2$beta.outcome[dat_pheno2$beta.exposure<0]
dat_pheno2$beta.exposure[dat_pheno2$beta.exposure<0]<--dat_pheno2$beta.exposure[dat_pheno2$beta.exposure<0]
mr_results <- mr(dat_pheno2, method_list=c("mr_ivw","mr_egger_regression","mr_weighted_median", "mr_weighted_mode"))
mr_results
```

```{r}
mr_scatter_plot(mr_results, dat_pheno2)
mr_forest_plot(mr_singlesnp(dat_pheno2))
```
