---
title: "Analysis over multiple generations"
author: "Clare Horscroft"
date: "30/03/2021"
output: html_document
---

What happens if a mutation stops being selected for? 20 generations of individuals are simulated in each scenario. In each scenario, the selection stops after 10 generations.

Load packages
```{r}
require(TwoSampleMR)
require(jsonlite)
```

Get file path
```{r}
config <- read_json("~/config.json")
mr_sel_path <- config$mr_sel_path
```

Use the Wald ratio to get the Biv estimate and standard error for each scenario and each generation
```{r}

b_mat<-matrix(NA,100,20)
se_mat<-matrix(NA,100,20)
for (s in 1:100){
  setwd(paste0(mr_sel_path,"/Simulations/children/multigen/selStop_sim_",s))
  for (g in 0:19){
    dat<-read.table(paste0("dat_",g,".txt"),header = TRUE)
    mod1<-summary(lm(dat$pheno~dat$geno))
    bgx<-mod1$coefficients[2,1]
    segx<-mod1$coefficients[2,2]
    mod2<-summary(glm(dat$children~dat$geno,family=poisson("log")))
    bgy<-mod2$coefficients[2,1]
    segy<-mod2$coefficients[2,2]
    mrw<-mr_wald_ratio(bgx,bgy,segx,segy)
    b_mat[s,g+1]<-mrw$b
    se_mat[s,g+1]<-mrw$se
  }
}
```

Compare selection coefficients the generation before selection ended and the generation after selection ended

```{r}
boxplot(b_mat,xlab="Generations since phenotype stopped being selected for",ylab="Estimated selection strength")
abline(h=0.05,lty=2,col="red")
abline(h=0,lty=2,col="blue")
```