---
title: "Proportion of variance explained"
author: "Clare Horscroft"
date: "19/08/2021"
output: html_document
---

Find the GWAS hits for the three variables of interest:

* Number of children
* Number of children fathered
* Number of live births

What is the proportion of variance explained by the SNPs identified in the GWAS?

Proportion of variance is calculated 3 ways: using the get_r_from_pn() and get_r_from_bsen() functions from the TwoSampleMR package, and the formula below from the linked paper: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0120758#sec017
Supplementary info gives this formula for proportion of variance explained:

$$
PVE = \frac{2\beta^2MAF(1-MAF)}{2\beta^2MAF(1-MAF)+2Nse^2MAF(1-MAF)}
$$

```{r setup}
library(rms2)
library(TwoSampleMR)
library(jsonlite)
# Get file path to project file 
config <- read_json("~/config.json")
mr_sel_path <- config$mr_sel_path
```

```{r initialise}
no_children <- rms2$new("ieu-b-4760")
no_live_births <- rms2$new("ukb-b-1209")
no_fathered <- rms2$new("ukb-b-2227")
```

```{r function}
prop_var_expl<-function(x){
  pve<-(2*(x$gwashits$beta^2)*x$gwashits$eaf*(1-x$gwashits$eaf))/(2*(x$gwashits$beta^2)*x$gwashits$eaf*(1-x$gwashits$eaf)+2*x$gwashits$n*(x$gwashits$se^2)*x$gwashits$eaf*(1-x$gwashits$eaf))
  return(list(
  pve=pve,
  pve_r_pn=get_r_from_pn(x$gwashits$p,x$gwashits$n)^2,
  pve_r_bsen=get_r_from_bsen(x$gwashits$beta,x$gwashits$se,x$gwashits$n)^2
  ))
}
```


## GWAS hits for number of children

```{r gwas_no_children}
no_children$extract_gwashits()
no_children$gwashits$n<-460654
no_children$gwashits
```

#### Proportion of variance explained

```{r pve_no_children}
no_children_pve<-prop_var_expl(no_children)
barplot(rbind(no_children_pve$pve,no_children_pve$pve_r_pn,no_children_pve$pve_r_bsen),beside = TRUE,names.arg = 1:length(no_children_pve$pve),legend.text = c("pve","pve_r_pn","pve_r_bsen"),main="Number of children",xlab="GWAS hit",ylab="Proportion of variance explained")
```

## GWAS hits for number of children fathered

```{r gwas_no_fathered}
no_fathered$extract_gwashits()
no_fathered$gwashits
```

#### Proportion of variance explained

```{r pve_no_fathered}
no_fathered_pve<-prop_var_expl(no_fathered)
barplot(rbind(no_fathered_pve$pve,no_fathered_pve$pve_r_pn,no_fathered_pve$pve_r_bsen),beside = TRUE,names.arg = 1:length(no_fathered_pve$pve),legend.text = c("pve","pve_r_pn","pve_r_bsen"),main="Number of children fathered",xlab="GWAS hit",ylab="Proportion of variance explained")
```

## GWAS hits for number of live births

```{r gwas_no_live_births}
no_live_births$extract_gwashits()
no_live_births$gwashits
```

#### Proportion of variance explained

```{r pve_no_live_births}
no_live_births_pve<-prop_var_expl(no_live_births)
barplot(rbind(no_live_births_pve$pve,no_live_births_pve$pve_r_pn,no_live_births_pve$pve_r_bsen),beside = TRUE,names.arg = 1:length(no_live_births_pve$pve),legend.text = c("pve","pve_r_pn","pve_r_bsen"),main="Number of live births",xlab="GWAS hit",ylab="Proportion of variance explained")
```

## Total variance explained

```{r}
barplot(cbind(sapply(no_children_pve,sum),
      sapply(no_fathered_pve,sum),
      sapply(no_live_births_pve,sum)),beside=TRUE,
      names.arg = c("no_children","no_fathered","no_live_births"),
      legend.text = c("pve","pve_r_pn","pve_r_bsen"),
      args.legend = list(x="top"),
      main="Total variance explained",
      ylab="Proportion of variance explained",
      ylim=c(0,0.002))
```

