---
title: Test simulation of children
author: Clare Horscroft
---

# Binary genotype

Set parameters
```{r}
set.seed(827441)

n<-100000
b<-0.2
```

Simulate parents' genotype and fitness
```{r}
parents_geno<-sample(c(0,1),n,TRUE)
parents_fitness<-1+b*(rnorm(n,0,1)+parents_geno)
tapply(parents_fitness,parents_geno,mean)
```

For each child choose two parents
```{r}
child_parent_1<-sample(1:n,n,replace=TRUE,prob=parents_fitness)
child_parent_2<-sample(1:n,n,replace=TRUE,prob=parents_fitness)
```

For each parent, count the number of children
```{r}
parents_no_children<-rep(0,n)
tabParents<-table(c(child_parent_1,child_parent_2))
parents_no_children[as.numeric(unlist(dimnames(tabParents)))]<-tabParents
barplot(table(parents_no_children))
```

Mean number of childen by genotype
```{r}
mean_by_geno<-tapply(parents_no_children,parents_geno,mean)
mean_by_geno
```
Difference is
```{r}
unname(mean_by_geno[2]-mean_by_geno[1])
```

geno = 1 divided by geno = 0 is
```{r}
unname(mean_by_geno[2]/mean_by_geno[1])
```

Create models
```{r}
mod1<-lm(parents_no_children~parents_geno)
mod2<-glm(parents_no_children~parents_geno,family=poisson)

#mod1
mod1

#mod2
#intercept
exp(mod2$coefficients[1])
#slope
exp(mod2$coefficients[1]) * (exp(mod2$coefficients[2]) - 1)
```

# Diploid genotype 

Set parameters
```{r}
set.seed(827441)

n<-100000
b<-0.2
```

Simulate parents' genotype and fitness
```{r}
parents_geno<-sample(0:2,n,TRUE)
parents_fitness<-1+b*(rnorm(n,0,1)+parents_geno)
tapply(parents_fitness,parents_geno,mean)
```

For each child choose two parents
```{r}
child_parent_1<-sample(1:n,n,replace=TRUE,prob=parents_fitness)
child_parent_2<-sample(1:n,n,replace=TRUE,prob=parents_fitness)
```

For each parent, count the number of children
```{r}
parents_no_children<-rep(0,n)
tabParents<-table(c(child_parent_1,child_parent_2))
parents_no_children[as.numeric(unlist(dimnames(tabParents)))]<-tabParents
barplot(table(parents_no_children))
```

Mean number of childen by genotype
```{r}
mean_by_geno<-tapply(parents_no_children,parents_geno,mean)
mean_by_geno
```
Difference is
```{r}
unname(c(mean_by_geno[2]-mean_by_geno[1],mean_by_geno[3]-mean_by_geno[2]))
unname(mean_by_geno[3]-mean_by_geno[1])
```

Ratios are
```{r}
unname(c(mean_by_geno[2]/mean_by_geno[1],mean_by_geno[3]/mean_by_geno[2]))
unname(mean_by_geno[3]/mean_by_geno[1])
```

Create models
```{r}
mod1<-lm(parents_no_children~parents_geno)
mod2<-glm(parents_no_children~parents_geno,family=poisson)
```
linear model
```{r}
mod1
```
generalised linear model
```{r}
#intercept
exp(mod2$coefficients[1])
#slope
exp(mod2$coefficients[1]) * (exp(mod2$coefficients[2]) - 1)
```

